Environmental Market Design for Large-Scale Marine Conservation

### 2019-11-20 docs/Markets_Conservation_revised_Nov14.tex
Updated MS for resubmission

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
It is commonly agreed that marine conservation should expand dramatically around the world. However, most countries have yet to implement large-scale no-take Marine Protected Areas (MPAs). When a country closes a large fraction of its waters to fishing, it stands to lose significant fishery revenue. While biodiversity and spillover fishing benefits may far exceed these losses, benefits from large-scale MPAs typically accrue to other countries or the high seas. To overcome this dilemma, we simulate and test an international fisheries management scheme with transferable fishing rights that incentivizes, rather than hinders, large-scale marine conservation. By combining a bioeconomic model of cross-country trading of fishing rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that transferable fishing rights and a biomass-based allocation rule are pivotal to incentivize conservation under this market-based setting. Our work focuses on the Vessel-Day Scheme, an environmental market employed by the Parties to the Nauru Agreement (a group of nine Pacific Island Nations) to manage their tuna fisheries, and where large-scale conservation interventions have taken place. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

@@

We are motivated by a real-world but relatively understudied institution called the Parties to the Nauru Agreement (PNA). The PNA is a coalition of nine Pacific Island Nations that collectively manages tuna purse seine fishing in its members' waters \cite{havice_2013,aqorau_2018} (Fig. \ref{fig:PNA_map}). These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to (mostly foreign) fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP \cite{ffa_2017b}. In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati, a PNA member, created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by January 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how an environmental market can be designed to incentivize conservation.

@@

Consistent with our model's prediction, after PIPA was implemented, displaced vessels relocated largely outside of Kiribati, and into other PNA countries' waters (Fig. \ref{fig:fishing_raster_diff}). Indeed, from 2014 to 2015, displaced vessels spent 2,115 fewer vessel-days (a 25\% reduction) in Kiribati, and 2,298 fewer vessel-days in PNA waters (a 17 \% reduction; Fig. \ref{fig:empirical}a-b). On the other hand, non-displaced vessels spent 4,656 additional days in Kiribati during 2015, and an additional 9,598 vessel-days in PNA waters. By 2018, we observe a net decrease of vessel-days within Kiribati, from 12,671 in 2014 to just 7,677 in 2018, with displaced vessels driving the decrease (Fig. \ref{fig:empirical}a). However, aggregate effort at the PNA-level remains relatively constant and we do not observe a ``fishing the line'' effect (Extended Data Figures 1 - 2). The reduction in effort in Kiribati and constant effort at the PNA-level suggest that trading facilitated redistribution of effort within PNA waters, just as the model predicts.

@@

As predicted by our model, the implementation of PIPA resulted in a decrease in fishing effort within Kiribati's water without large revenue losses. Kiribati's reported revenue increased from US\$127.3 million in 2014 to US\$148.8 million in 2015, before decreasing to US\$118.3 million in 2016 (Fig. \ref{fig:empirical}c). The increase and subsequent decrease in revenues matches the vessel-day patterns observed for Kiribati from 2014 to 2016 (Fig. \ref{fig:empirical}d). But, critically, the drop in revenue in 2016 (20\%) is smaller than the drop in vessel-days (35\%). This supports a key prediction of our model: with trading, the relative revenue drop will be smaller than the relative effort drop but, without trading, the opposite relationship would hold (Supplementary Figure 4). At the PNA level, total revenues showed a net increase of \$64.7 and \$28 million USD for 2015 and 2016 respectively (Supplementary Figure 6), despite the PIPA closure.

@@

Our findings may help inform management and implementation of existing and upcoming MPAs in the PNA. In January 2020, Palau will close nearly 80\% of its EEZ to commercial fishing to create the Palau National Marine Sanctuary (PNMS): the 14\textsuperscript{th} largest protected area in the world (Fig. \ref{fig:PNA_map}). Vessel-tracking data (2012 to 2018) shows that, on average, the proposed PNMS boundaries have historically contained 86 $\pm$ 5.30\% ($\pm$ 1SD) of longline vessel-days (non-tradable) and 91.3 $\pm$ 5.03\% ($\pm$ 1SD) of purse seine vessel-days (tradable) in Palauan waters (Extended Data Figures 3 - 4).

@@

We begin by filtering the data to keep only positions labeled as fishing events. We then create a gridded version of the data for each year and group (\emph{i.e.} displaced and non-displaced) by binning the coordinates to a 1-degree grid and summing all fishing hours for a given grid cell. We use a 1-degree grid as a reasonable compromise between higher resolutions that would result in a more granular but noisy footprint, and the simple estimation of vessel-days at the EEZ-level (as in Extended Data Figure 1, but represented spatially). This process results in 14 gridded datasets of fishing hours (7 years, for two groups). For each group of vessels, we then calculate the average fishing hours before (2012 - 2014, inclusive) and after (2015 - 2018, inclusive) the implementation of PIPA, resulting in 4 datasets of mean fishing effort (before and after for displaced and non-displaced vessels). We then calculate the change in effort allocation between these two periods (after minus before) for each group, and normalize the value of each grid cell by diving it by the largest within-group absolute change:

@@

We obtained information on revenues from the Pacific Islands Forum Fisheries Agency \emph{Tuna Development Indicators 2016} report.  Specifically, we use data compiled by the Pacific Islands Forum Fisheries Agency (FFA \cite{ffa_2017}) where annual revenues from license fees (for VDS and other access programs) are reported for each country (2008 - 2016; Fig. \ref{fig:empirical}c; Supplementary Figures 6 - 7). For countries in the PNA, these revenues show a combination of vessel-day license fees as well as joint-venture operations.

@@

%\bibliography{references}
%\bibliographystyle{nature}

\begin{thebibliography}{10}

\bibitem{oleary_2016}
O'Leary, B.~C., Winther-Janson, M., Bainbridge, J.~M., Aitken, J., Hawkins,
  J.~P., and Roberts, C.~M.
\newblock {\em Conservation letters}{ \bf 9}(6), 398--404 nov  (2016).

\bibitem{dinerstein_2019}
Dinerstein, E., Vynne, C., Sala, E., Joshi, A.~R., Fernando, S., Lovejoy,
  T.~E., Mayorga, J., Olson, D., Asner, G.~P., Baillie, J. E.~M., Burgess,
  N.~D., Burkart, K., Noss, R.~F., Zhang, Y.~P., Baccini, A., Birch, T., Hahn,
  N., Joppa, L.~N., and Wikramanayake, E.
\newblock {\em Sci. Adv.}{ \bf 5}(4), eaaw2869 apr  (2019).

\bibitem{sala_2017}
Sala, E. and Giakoumi, S.
\newblock {\em {ICES} Journal of Marine Science}{ \bf } (2017).

\bibitem{sala_2018}
Sala, E., Lubchenco, J., Grorud-Colvert, K., Novelli, C., Roberts, C., and
  Sumaila, U.~R.
\newblock {\em Marine Policy}{ \bf 91}(2), 11--13 may  (2018).

\bibitem{mcleod_2019}
Mcleod, E., Bruton-Adams, M., Förster, J., Franco, C., Gaines, G., Gorong, B.,
  James, R., Posing-Kulwaum, G., Tara, M., and Terk, E.
\newblock {\em Frontiers in Marine Science}{ \bf } (2019).

\bibitem{ramesh_2019}
Ramesh, N., Rising, J.~A., and Oremus, K.~L.
\newblock {\em Science}{ \bf 364}(6446), 1192--1196 jun  (2019).

\bibitem{hernndez_2019}
Hernández, C.~M., Witting, J., Willis, C., Thorrold, S.~R., Llopiz, J.~K., and
  Rotjan, R.~D.
\newblock {\em Scientific reports}{ \bf 9}(1), 10772 dec  (2019).

\bibitem{agardy_2018}
Agardy, T.
\newblock {\em {ICES} Journal of Marine Science}{ \bf 75}(3), 1183--1185 may
  (2018).

\bibitem{havice_2010}
Havice, E.
\newblock {\em Marine Policy}{ \bf 34}(5), 979--987 sep  (2010).

\bibitem{havice_2013}
Havice, E.
\newblock {\em Marine Policy}{ \bf 42}, 259--267 nov  (2013).

\bibitem{aqorau_2018}
Aqorau, T., Bell, J., and Kittinger, J.~N.
\newblock {\em Science}{ \bf 361}(6408), 1208.2--1209 sep  (2018).

\bibitem{ffa_2017b}
FFA.
\newblock Technical report, Pacific Islands Forum Fisheries Agency,  (2017).

\bibitem{libecap_1989}
Libecap, G.~D.
\newblock {\em Journal of Institutional and Theoretical Economics}{ \bf
  145}(1), 6--24 mar  (1989).

\bibitem{mccauley_2016}
{McCauley}, D.~J., Woods, P., Sullivan, B., Bergman, B., Jablonicky, C., Roan,
  A., Hirshfield, M., Boerder, K., and Worm, B.
\newblock {\em Science}{ \bf 351}(6278), 1148--1150 mar  (2016).

\bibitem{mcdermott_2018}
{McDermott}, G.~R., Meng, K.~C., {McDonald}, G.~G., and Costello, C.~J.
\newblock {\em Proc Natl Acad Sci {USA}}{ \bf } aug  (2018).

\bibitem{kroodsma_2018}
Kroodsma, D.~A., Mayorga, J., Hochberg, T., Miller, N.~A., Boerder, K.,
  Ferretti, F., Wilson, A., Bergman, B., White, T.~D., Block, B.~A., Woods, P.,
  Sullivan, B., Costello, C., and Worm, B.
\newblock {\em Science}{ \bf 359}(6378), 904--908 feb  (2018).

\bibitem{ffa_2017}
Terawasi, P. and Reid, C.
\newblock Technical report, Pacific Islands Forum Fisheries Agency, Honaira,
  Solomon Islands,  (2017).

\bibitem{hanich2018unraveling}
Hanich, Q., Rotjan, R., Aqorau, T., Bailey, M., Campbell, B., Gray, N., Gruby,
  R., Hampton, J., Ota, Y., Parris, H., et~al.
\newblock {\em Proceedings of the National Academy of Sciences}{ \bf 115}(52),
  E12122--E12123 (2018).

\bibitem{ferraro2018causal}
Ferraro, P.~J., Sanchirico, J.~N., and Smith, M.~D.
\newblock {\em Proceedings of the National Academy of Sciences}{ \bf },
  201805563 (2018).

\bibitem{crespo_2019}
Crespo, G.~O., Dunn, D.~C., Gianni, M., Gjerde, K., Wright, G., and Halpin,
  P.~N.
\newblock {\em Nature Ecology \& Evolution}{ \bf 3}(9), 1273--1276 sep  (2019).

\bibitem{rosegrant_1994}
Rosegrant, M.~W. and Binswanger, H.~P.
\newblock {\em World development}{ \bf 22}(11), 1613--1625 nov  (1994).

\bibitem{hutton_2003}
Hutton, J.~M. and Leader-Williams, N.
\newblock {\em Oryx}{ \bf 37}(02) apr  (2003).

\bibitem{sandbrook_2019}
Sandbrook, C., Fisher, J., Holmes, G., Luque-Lora, R., and Keane, A.
\newblock {\em Nat Sustain}{ \bf 2}(4), 316--323 apr  (2019).

\bibitem{roberts_2017}
Roberts, C.~M., O'Leary, B.~C., {McCauley}, D.~J., Cury, P.~M., Duarte, C.~M.,
  Lubchenco, J., Pauly, D., S√°enz-Arroyo, A., Sumaila, U.~R., Wilson,
  R.~W., Worm, B., and Castilla, J.~C.
\newblock {\em Proc Natl Acad Sci {USA}}{ \bf 114}(24), 6167--6175 jun  (2017).

\bibitem{ban_2019}
Ban, N.~C., Gurney, G.~G., Marshall, N.~A., Whitney, C.~K., Mills, M., Gelcich,
  S., Bennett, N.~J., Meehan, M.~C., Butler, C., Ban, S., Tran, T.~C., Cox,
  M.~E., and Breslow, S.~J.
\newblock {\em Nature Sustainability}{ \bf 2}(6), 524--532 jun  (2019).

\bibitem{costello_2016}
Costello, C., Ovando, D., Clavelle, T., Strauss, C.~K., Hilborn, R., Melnychuk,
  M.~C., Branch, T.~A., Gaines, S.~D., Szuwalski, C.~S., Cabral, R.~B., Rader,
  D.~N., and Leland, A.
\newblock {\em Proceedings of the National Academy of Sciences of the United
  States of America}{ \bf 113}(18), 5125--5129 may  (2016).

\bibitem{Hagrannsoknir2014}
Hagrannsoknir.
\newblock Technical report, PNA Office, available at: www.pnatuna.com/sites/
  default/files/Independent Review of the PNA Purse Seine VDS- Final Report_0.pdf,  (2014).

\bibitem{pna_2016}
PNA.
\newblock Technical report, Parties to the Nauru Agreement,  oct  (2016).

\bibitem{rcore_2019}
{R Core Team}.
\newblock {\em R: A Language and Environment for Statistical Computing}.
\newblock R Foundation for Statistical Computing, Vienna, Austria,  (2019).

\bibitem{yeeting2018stabilising}
Yeeting, A.~D., Weikard, H.-P., Bailey, M., Ram-Bidesi, V., and Bush, S.~R.
\newblock {\em Regional Environmental Change}{ \bf 18}(3), 885--897 (2018).

\end{thebibliography}

>>>>

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
It is commonly agreed that marine conservation should expand dramatically around the world. However, most countries have yet to implement large-scale no-take Marine Protected Areas (MPAs). When a country closes a large fraction of its waters to fishing, it stands to lose significant fishery revenue. While biodiversity and spillover fishing benefits may far exceed these losses, benefits from large-scale MPAs typically accrue to other countries or the high seas. To overcome this dilemma, we simulate and test an international fisheries management scheme with transferable fishing rights that incentivizes, rather than hinders, large-scale marine conservation. By combining a bioeconomic model of cross-country trading of fishing rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that cross-country transferable fishing rights and a biomass-based allocation rule are pivotal to incentivize conservation under this market-based setting. Our work focuses on the Vessel-Day Scheme, an environmental market employed by the Parties to the Nauru Agreement (a group of nine Pacific Island Nations) to manage their tuna fisheries, and where large-scale conservation interventions have taken place. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

@@

We are motivated by a real-world but relatively understudied institution called the Parties to the Nauru Agreement (PNA). The PNA is a coalition of nine Pacific Island Nations that collectively manages tuna purse seine fishing in its members' waters \cite{havice_2013,aqorau_2018} (Fig. \ref{fig:PNA_map}). These waters account for 14.5 million km\textsuperscript{2} (an area four times larger than the continental US), and over 60\% of skipjack tuna catch in the Western Central Pacific \cite{havice_2013}. The PNA manages this tuna purse seine fishery using a vessel-day scheme (VDS) where total annual fishing effort is capped at around 45,000 vessel-days. Vessel-days are allocated across the nine nations, which then lease them to (mostly foreign) fishing vessels. A vessel-day grants a fishing vessel the right to fish for 24 hours within one of the nine Exclusive Economic Zones (EEZs) within the PNA. Member nations derive enormous benefits from leasing these fishing rights to foreign fleets, in some cases exceeding half of a country's GDP \cite{ffa_2017b}. In addition to highly productive tuna fisheries, the PNA waters provide a wealth of ecosystem benefits, hence the focus on large-scale conservation efforts in the region \cite{mcleod_2019}. In 2015, Kiribati, a PNA member, created the Phoenix Islands Protected Area, one of the largest protected areas on earth (408,250 km\textsuperscript{2}), and Palau will close 80\% of its national waters to fishing by January 2020. We draw from the PNA's market-based approach to managing tuna fisheries and build on it to show how a market can be designed to incentivize conservation.

@@

Consistent with our model's prediction, after PIPA was implemented, displaced vessels relocated largely outside of Kiribati, and into other PNA countries' waters (Fig. \ref{fig:fishing_raster_diff}). Indeed, from 2014 to 2015, displaced vessels spent 2,115 fewer vessel-days (a 25\% reduction) in Kiribati, and 2,298 fewer vessel-days in PNA waters (a 17 \% reduction; Fig. \ref{fig:empirical}a-b). On the other hand, non-displaced vessels spent 4,656 additional days in Kiribati during 2015, and an additional 9,598 vessel-days in PNA waters. By 2018, we observe a net decrease of vessel-days within Kiribati, from 12,671 in 2014 to just 7,677 in 2018, with displaced vessels driving the decrease (Fig. \ref{fig:empirical}a). However, aggregate effort at the PNA-level remains relatively constant and we do not observe a ``fishing the line'' effect (Supplementary Figures 6 - 7). The reduction in effort in Kiribati and constant effort at the PNA-level suggest that trading facilitated redistribution of effort within PNA waters, just as the model predicts.

@@

As predicted by our model, the implementation of PIPA resulted in a decrease in fishing effort within Kiribati's water without large revenue losses. Kiribati's reported revenue increased from US\$127.3 million in 2014 to US\$148.8 million in 2015, before decreasing to US\$118.3 million in 2016 (Fig. \ref{fig:empirical}c). The increase and subsequent decrease in revenues matches the vessel-day patterns observed for Kiribati from 2014 to 2016 (Fig. \ref{fig:empirical}d). But, critically, the drop in revenue in 2016 (20\%) is smaller than the drop in vessel-days (35\%). This supports a key prediction of our model: with trading, the relative revenue drop will be smaller than the relative effort drop but, without trading, the opposite relationship would hold (Supplementary Figure 4). At the PNA level, total revenues showed a net increase of \$64.7 and \$28 million USD for 2015 and 2016 respectively (Supplementary Figure 8), despite the PIPA closure.

@@

Our findings may help inform management and implementation of existing and upcoming MPAs in the PNA. In January 2020, Palau will close nearly 80\% of its EEZ to commercial fishing to create the Palau National Marine Sanctuary (PNMS): the 14\textsuperscript{th} largest protected area in the world (Fig. \ref{fig:PNA_map}). Vessel-tracking data (2012 to 2018) shows that, on average, the proposed PNMS boundaries have historically contained 86 $\pm$ 5.30\% ($\pm$ 1SD) of longline vessel-days (non-tradable) and 91.3 $\pm$ 5.03\% ($\pm$ 1SD) of purse seine vessel-days (tradable) in Palauan waters (Supplementary Figures 10-11).

@@

We begin by filtering the data to keep only positions labeled as fishing events. We then create a gridded version of the data for each year and group (\emph{i.e.} displaced and non-displaced) by binning the coordinates to a 1-degree grid and summing all fishing hours for a given grid cell. We use a 1-degree grid as a reasonable compromise between higher resolutions that would result in a more granular but noisy footprint, and the simple estimation of vessel-days at the EEZ-level (as in Supplementary Figure 6, but represented spatially). This process results in 14 gridded datasets of fishing hours (7 years, for two groups). For each group of vessels, we then calculate the average fishing hours before (2012 - 2014, inclusive) and after (2015 - 2018, inclusive) the implementation of PIPA, resulting in 4 datasets of mean fishing effort (before and after for displaced and non-displaced vessels). We then calculate the change in effort allocation between these two periods (after minus before) for each group, and normalize the value of each grid cell by diving it by the largest within-group absolute change:

@@

We obtained information on revenues from the Pacific Islands Forum Fisheries Agency \emph{Tuna Development Indicators 2016} report.  Specifically, we use data compiled by the Pacific Islands Forum Fisheries Agency (FFA \cite{ffa_2017}) where annual revenues from license fees (for VDS and other access programs) are reported for each country (2008 - 2016; Fig. \ref{fig:empirical}c; Supplementary Figures 8 - 9). For countries in the PNA, these revenues show a combination of vessel-day license fees as well as joint-venture operations.

@@

\bibliography{references}
\bibliographystyle{nature}

### 2019-11-20 docs/old_ms/Markets_Conservation_revised_Nov13.tex
Updated MS for resubmission

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
It is commonly agreed that marine conservation should expand dramatically around the world. However, most countries have yet to implement large-scale no-take Marine Protected Areas (MPAs). When a country closes a large fraction of its waters to fishing, it stands to lose significant fishery revenue. While biodiversity and spillover fishing benefits may far exceed these loses, benefits from large-scale MPAs typically accrue to other countries or the high seas. To overcome this dilemma, we design and test an international fisheries management scheme with transferable fishing rights that incentivizes, rather than hinders, large-scale marine conservation. By combining a model of cross-country trading of fishing rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that cross-country transferable fishing rights and a biomass-based allocation rule are pivotal to incentivize conservation. Our work focuses on the Vessel-Day Scheme, an environmental market employed by the Parties to the Nauru Agreement (a group of nine Pacific Island Nations) to manage their tuna fisheries, and where large-scale conservation interventions have taken place. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}

>>>>

% Place your abstract within the special {sciabstract} environment.
\begin{abstract}
It is commonly agreed that marine conservation should expand dramatically around the world. However, most countries have yet to implement large-scale no-take Marine Protected Areas (MPAs). When a country closes a large fraction of its waters to fishing, it stands to lose significant fishery revenue. While biodiversity and spillover fishing benefits may far exceed these losses, benefits from large-scale MPAs typically accrue to other countries or the high seas. To overcome this dilemma, we design and test an international fisheries management scheme with transferable fishing rights that incentivizes, rather than hinders, large-scale marine conservation. By combining a model of cross-country trading of fishing rights with vessel-level tracking data before and after a large-scale conservation action is implemented, we show that cross-country transferable fishing rights and a biomass-based allocation rule are pivotal to incentivize conservation. Our work focuses on the Vessel-Day Scheme, an environmental market employed by the Parties to the Nauru Agreement (a group of nine Pacific Island Nations) to manage their tuna fisheries, and where large-scale conservation interventions have taken place. Overall, these results provide a template for how to incentivize countries to engage in large-scale marine conservation within a market-based setting.
\end{abstract}
